---
output: 
  github_document:
    pandoc_args: --webtex

---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  cache = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# pollagg

<!-- badges: start -->
<!-- badges: end -->

`pollagg` is an out-of-the-box poll aggregator for researchers who want to understand trends in public opinion from multiple polls or surveys taken over time.


## Installation

You can install the development version of `pollagg` with

``` r
devtools::install_github("alexpavlakis/pollagg")
```
# Examples

The poll aggregation function `yapa` (yet another poll aggregator) takes three main arguments:

1. `y` is a matrix of counts for responses (columns) for each poll or survey (rows);
2. `n` is a vector of the total counts for each poll or survey;
3. `dates` is a vector of dates (can be dates or numeric) for when each poll or survey was conducted.

An additional argument `all_dates` is available if you want to produce specific estimates for dates other than survey dates.  For example, if you conduct surveys on the first of each month but want to estimate hypothetical results if you had conducted them on the 15th instead.

`yapa` returns a list with:

* `trend` a `data.frame` of the underlying trend and it's 90% uncertainty interval;
* `pct` a `data.frame` of the smoothed poll averages and their 90% uncertainty intervals;
* `delta` a `data.frame` of the estimated changes in trend and their 90% uncertainty intervals;
* `polls` a `data.frame` of the raw polling data on which the model is built;
* And other elements from the model arguments and results.

```{r obama_mccain_data}
library(pollagg)
suppressMessages(library(dplyr))
options(mc.cores = parallel::detectCores())
rstan::rstan_options(auto_write = TRUE)

# General election polls - Obama vs McCain 2008
head(obama_mccain_polls)
```

In this example, we aggregate all polls conducted of the 2008 presidential race between January 1st and election day 2008.  This is a very noisy series but the model is able to pick up on the underlying trend as well as clear breaks, such as the collapse of Lehman Brothers on September 15th, which pushed the popular vote toward Barack Obama.


```{r obama_mccain_model, results = F}

y <- select(obama_mccain_polls, `McCain (R)`, `Obama (D)`)
n <- obama_mccain_polls$n
dates <- obama_mccain_polls$poll_end

fit_polls <- yapa(y = y, n = n, dates = dates)

```


```{r obama_mccain_res}
plot(fit_polls, size = 0.5) + 
  ylim(0.3, 0.6) +
  scale_x_date(date_breaks = '1 month') +
  scale_fill_manual(values = c('red', 'blue')) +
  scale_color_manual(values = c('red', 'blue')) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.subtitle = element_text(size = 9),
        legend.position = 'bottom') +
  labs(x = NULL, y = NULL, title = 'General Election: McCain vs. Obama',
       subtitle = 'polls (faint dots), modeled polls (bold dots), trend (line), and trend uncertainty (faded ribbon)', col = NULL, fill = NULL)
```

The model can also estimate clear trends from large-sample, precise surveys, such as trends in responses to the General Social Survey question regarding the legality of marijuana between 1973 and 2018.


```{r gss_data}
head(grass_gss)
```

```{r gss_model, results = F}

y <- select(grass_gss, legal, illegal)
n <- grass_gss$n
dates <- grass_gss$year

fit_gss <- yapa(y = y, n = n, dates = dates)
```

```{r gss_res}

plot(fit_gss) + 
  ylim(0, 1) +
  scale_fill_manual(values = c('darkgreen', 'darkgrey')) +
  scale_color_manual(values = c('darkgreen', 'darkgrey')) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.subtitle = element_text(size = 9),
        legend.position = 'bottom') +
  labs(x = NULL, y = NULL, title = 'Should marijuana be made legal?',
       subtitle = 'survey results (faint dots), modeled survey results (bold dots), trend (line), and trend uncertainty (faded ribbon)', col = NULL, fill = NULL)
```



# Methodology

The underlying framework is a hierarchical, nonlinear model which partially pools poll averages ($\theta$) for each option ($o$) in each poll ($p$) toward their trend ($\mu$) on each day ($d$).  The trend is estimated from a baseline rate ($\alpha$) and daily changes ($\delta$), which naturally estimates nonlinear trends.

The model is fit with [Stan](www.https://mc-stan.org/).


$$
\begin{align*}
y_{p, o} &\sim Binomial(n_p, \theta_{p, o}) \\
\theta_{p, o} &\sim Normal(\mu_{d(p), o}, \sigma^\theta) \\
\mu_{d, o} &= \sum_{t < d} \delta_{t, o} + \alpha_o \\
\end{align*}
$$
